---
title: "Homework 7"
author: "Nishka Rastogi"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, eval = F, warning = F, message = F)
library(tidyverse)
library(MatchIt)
armfold <- read.csv("armfold.csv")
turnout <- read.csv("turnout.csv")

```

Name: Nishka Rastogi

UT EID: nr26674

GitHub Link: <https://github.com/rastogin1216/SDS315>

## Problem 1: Armfolding

### Part A

```{r}
males <- armfold |>
  filter(Sex == "Male") 
male_size <- males |>
  nrow()

male_sample_prop <- (males |>
  filter(LonR_fold == "1") |>
  nrow()) / male_size


females <- armfold |>
  filter(Sex == "Female") 
female_size <- females |>
  nrow()

females_sample_prop <- (females |>
  filter(LonR_fold == "1") |>
  nrow()) / female_size

```

There is a total of 106 males and 111 females in this sample. The sample proportion of males who folded their left arm on top is 50/106 (about 0.472), and the sample proportion of females who folded their left arm on top is 47/111 (about 0.423).

### Part B

```{r}
difference_prop <- male_sample_prop - females_sample_prop


```

The observed difference in proportions between males and females is about 0.048.

### Part C

```{r}
stand_error <- sqrt((male_sample_prop * (1 - male_sample_prop)) / male_size + (females_sample_prop * (1 - females_sample_prop)) / female_size)

prop.test(x = c(50,47), n = c(male_size,female_size),conf.level = 0.95,correct = F)

```

SE(phat1 - phat2) = sqrt((phat1(1-phat1))/N1) + (phat2(1-phat2))/N2))

phat1 = 0.472

N1 = 106

phat2 = 0.423

N2 = 111

I used z = 1.96 for a 95% interval because 95% of the probability in a standard normal distribution falls within about 1.96 standard deviations of the mean. Thus, this represents the critical value, where each tail of the distribution contains about 2.5% of the probability.

The standard error I calculated is about 0.067.

### Part D

If we were to estimate the true difference in proportions of left-over-right arm folding between male and female undergraduate students at this university, the estimand would fall between -0.084 and 0.180, with 95% confidence.

### Part E

The standard error represents the typical variation of the difference in sample proportions between males and females compared to its expected value under repeated sampling. It measures how much the observed difference in arm-folding proportions between sexes would vary from sample to sample due to sampling variability.

### Part F

The sampling distribution in this context refers to the distribution of the difference in sample proportions between males and females that we would obtain if we repeatedly randomly sampled from the population and computed this difference each time. Thus, the sample proportions (and hence the difference in sample proportions of left-over-right arm folding between sexes) will vary from sample to sample due to the differences in individuals that comprise each sample. Additionally, the sample sizes of males and females stay fixed.

### Part G

The Central Limit Theorem justifies using a normal distribution (provided that the sample sizes are sufficiently large) to approximate the sampling distribution of the difference in sample proportions. The difference in sample proportions can be interpreted as an aggregation of nudges, where randomness causes small fluctuations in both directions; these fluctuations balance out over many samples, producing a normal distribution for the sample distribution of the difference in sample proportions.

### Part H

If someone tried to claim that there is no sex difference in arm folding based on a 95% confidence interval for the difference in proportions, [-0.01,0.30]. It is true that this confidence interval contains 0, so it is possible that there is no true difference in proportions of left-over-right arm folding between sexes, thus making the result statistically insignificant. However, this claim is not definitively proven nor refuted because our confidence interval also suggests that there could be a difference of up to 30%, so we do not have strong enough evidence to make a definitive conclusion like the claim made.

### Part I

The confidence interval itself would be different across samples because the sample proportions would be different from sample to sample due to random sampling variability. Nonetheless, if we were to repeat this experiment many times and compute a confidence interval for each sample, about 95% of the intervals would capture the true difference in proportions according to the coverage principle.

## Problem 2: Get out the Vote

### Part A

```{r}
gotv <- turnout |>
  filter(GOTV_call == 1) 

sum(gotv$voted1998) / nrow(gotv)

not_gotv <- turnout |>
  filter(GOTV_call == 0)
sum(not_gotv$voted1998) / nrow(not_gotv)

prop.test(x = c(160,4701), n = c(nrow(gotv),nrow(not_gotv)),conf.level = 0.95,correct = F)
```

The proportion of those receiving a GOTV call who voted in 1998 is about 0.648. The proportion of those not receiving a GOTV call who voted in 1998 is about 0.444. The true difference in proportions of voting in 1998 for those who received a GOTV call versus those who didn't is between about 0.143 and about 0.264, with 95% confidence.

### Part B

```{r,eval = T}
turnout |>
  ggplot() + geom_boxplot(aes(x = AGE))+
  facet_wrap(~ GOTV_call, labeller = label_both) +
  theme_minimal() +
  labs(title = "Distribution of Age by GOTV Call Group",
       x = "Age",
       y = "Count")

turnout |>
  ggplot() + geom_bar(aes(x = as.factor(MAJORPTY),fill = as.factor(MAJORPTY)))+
  facet_wrap(~ GOTV_call, labeller = label_both) +
  theme_minimal() +
  labs(title = "Barplot of Party by GOTV Call Group",
       x = "Major Party (0 = No, 1 = Yes)",
       y = "Count") +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "darkblue"),
                    name = "Major Party")


turnout |>
  ggplot() + geom_bar(aes(x = as.factor(voted1996),fill = as.factor(voted1996)))+
  facet_wrap(~ GOTV_call, labeller = label_both) +
  theme_minimal() +
  labs(title = "Barplot of Voted in 1996 by GOTV Call Group",
       x = "Voted in 1996 (0 = No, 1 = Yes)",
       y = "Count") + 
  scale_fill_manual(values = c("0" = "lightblue", "1" = "darkblue"),
                    name = "Voted in 1996")

```

```{r}


t.test(AGE ~ GOTV_call, data = turnout)


prop.test(x = c(sum(gotv$voted1996),sum(not_gotv$voted1996)), n = c(nrow(gotv),nrow(not_gotv)),conf.level = 0.95,correct = F)

prop.test(x = c(sum(gotv$MAJORPTY),sum(not_gotv$MAJORPTY)), n = c(nrow(gotv),nrow(not_gotv)),conf.level = 0.95,correct = F)
```

The true difference in mean ages between those who did and didn't receive GOTV calls is between about 6.370 and 11.395, with 95% confidence.

The true difference in proportions of major party voters between those who did and didn't receive GOTV calls is between about 0.006 and 0.107, with 95% confidence.

The true difference in proportions of those who did and didn't vote in 1996 between those who did and didn't receive GOTV calls is between about 0.125 and 0.239, with 95% confidence.

### Part C

```{r, eval = T}

turnout_matched <- matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY, data = turnout, ratio = 5) 
matched_data <- match.data(turnout_matched)

matched_data |>
  group_by(GOTV_call) |>
  summarise(`Mean Age` = mean(AGE, na.rm = TRUE))

tab1 <- table(as.factor(matched_data$voted1996), as.factor(matched_data$GOTV_call))
dimnames(tab1) <- list(
  "Voted in 1996" = c("No", "Yes"),
  "GOTV Call" = c("No Call", "Called")
)
prop.table(tab1, margin = 2)
tab2 <- table(as.factor(matched_data$MAJORPTY), as.factor(matched_data$GOTV_call))
dimnames(tab2) <- list(
  "Major Party" = c("No", "Yes"),
  "GOTV Call" = c("No Call", "Called")
)
prop.table(tab2, margin = 2)
```

```{r}

gotv2 <- matched_data |>
  filter(GOTV_call == 1) 

sum(gotv2$voted1998) / nrow(gotv2)

not_gotv2 <- matched_data |>
  filter(GOTV_call == 0)
sum(not_gotv2$voted1998) / nrow(not_gotv2)

prop.test(x = c(160,703), n = c(nrow(gotv2),nrow(not_gotv2)),correct = F)


```

Using the matched dataset, the proportion of those receiving a GOTV call who voted in 1998 is about 0.648. The proportion of those not receiving a GOTV call who voted in 1998 is about 0.569. The true difference in proportions of voting in 1998 for those who received a GOTV call versus those who didn't is between about 0.013 and about 0.144, with 95% confidence.

Thus, I can conclude that there is convincing evidence that the GOTV call had a positive and statistically significant effect on voter turnout in the 1998 election. However, this effect may be practically trivial.
