---
title: "Homework 8"
author: "Nishka Rastogi"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, eval = F, warning = F, message = F)
library(tidyverse)
library(mosaic)
covid <- read.csv("covid.csv")
creatinine <- read.csv("creatinine.csv")
milk <- read.csv("milk.csv")
```

Name: Nishka Rastogi

UT EID: nr26674

GitHub Link: <https://github.com/rastogin1216/SDS315>

# Problem 1

```{r}

model_hr = lm(creatclear ~ age, data=creatinine)
coef(model_hr)
```

## Part A

The expected creatinine clearance rate for a 55-year-old patient would be about **113.72 mL/minute**.

I fitted a linear regression model using the data and outputted the intercept and slope of the model.

Our fitted regression model is:

y-hat = 147.81 - 0.62(x)

where y-hat is the predicted creatinine clearance rate and x is the age.

Using this model, I was able to predict the expected creatinine clearance rate for a 55-year-old patient by plugging in 55 as x in our fitting regression model.

y-hat = 147.81 - 0.62(55)

y-hat = 113.72

## Part B

The slope of **-0.62 ml/minute per year** represents the relationship between creatinine clearance rate and age. Therefore, a one-year change in age is associated with a -0.62 ml/minute change in creatinine clearance rate, on average. I determined this by extracting the slope from the generated linear regression model.

y-hat = 147.81 **-** **0.62**(x)

## Part C

The 40-year-old has a creatinine clearance rate that is healthier for their age.

I determined this by calculating the residual for each patient, using their predicted creatinine clearance rates from the linear regression model.

**40-year-old:**

Residual = Actual Value - Predicted Value

Residual = 135 - (147.81 - 0.62(40))

Residual = 11.98

**60-year-old:**

Residual = Actual Value - Predicted Value

Residual = 112- (147.81 - 0.62(60))

Residual = 1.38

Because the 40-year-old has a higher residual than the 60-year-old, the 40-year-old has a creatinine clearance rate that is healthier for their age.

# Problem 2

```{r}
italy <- covid |>
  filter(country == "Italy")

spain <- covid |>
  filter(country == "Spain")

italy_boot = do(10000)*lm(log(deaths) ~ days_since_first_death, data=resample(italy))
confint(italy_boot, level = 0.95)

spain_boot = do(10000)*lm(log(deaths) ~ days_since_first_death, data=resample(spain))
confint(spain_boot, level = 0.95)
```

## Part A

The estimated growth rate for Italy, with 95% confidence, is between 0.159 and 0.209. The doubling time is between about 335.2 and 438.9 days, with 95% confidence.

## Part B

\
The estimated growth rate for Spain is between 0.235 and 0.317, with 95% confidence. The doubling time is between about 220.8 and 297.9 days, with 95% confidence.

## Part C

```{r, eval = T}
ggplot(data = covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1) +
  labs(
    title = "Daily COVID-19 Deaths Since First Death",
    x = "Days Since First Death",
    y = "Number of Deaths",
    color = "Country"
  ) +
  theme_minimal()
```

# Problem 3

```{r}
lm_milk = lm(log(sales) ~ log(price), data=milk)
coef(lm_milk)
milk_boot = do(10000)*lm(log(sales) ~ log(price), data=resample(milk))
confint(milk_boot, level = 0.95)


```

The estimated price elasticity of demand for milk is about -1.62, in light of the data; thus, when milk prices increase by 1%, consumers want to buy about 1.62% less of it, on average.

The true price elasticity of demand for milk is between about -1.77 and -1.45, with 95% confidence. I estimated this quantity by creating many fitted regression models on a log-to-log scale. I used the fitted slope of each regression model to determine the elasticity.
